
White Wine Quality Between Harmfulness and Benefit By Exploring Physicochemical Properties Data
===============

# Introduction
   The aim of  this project is describe the physicochemical properties to try 
specifying which of them affect white wine quality through exploratory data 
analysis of a dataset containing attributes for 4898 white wine variants.
   The dataset is related to **"White Wine"** of the Portuguese "Vinho Verde" 
wine. For more details, consult: http://www.vinhoverde.pt/en/ or the reference 
[Cortez et al., 2009]. Due to privacy and logistic issues, only physicochemical 
(inputs) and sensory (the output) variables are available (e.g. there is no data 
about grape types, wine brand, wine selling price, etc.).

- Number of Instances (obs.): 4898. 
- Number of Attributes (variables): 11 + output attribute + serial No (x var.)

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Loading the ggplot graphics package and the others
library(ggplot2)
library(ggthemes)
library(GGally)
library(corrplot)
library(gridExtra)
```

```{r echo=FALSE, Load_the_Data}
# Load the Data
white_wine <- read.csv("wineQualityWhites.csv", sep = ",")
```

# Univariate Data Exploring and Plots Section

```{r echo=FALSE, dimensions}
dim(white_wine)
```

Our dataset consists of 13 variables, with 4898 observations.

```{r echo=FALSE, string}
str(white_wine)
```

All the variables are numerical which quantify the chemical properties of 
the wine, except 'quality' variable is integer. At least 3 wine experts rated 
the quality of each wine, providing a rating between 0 (very bad) and 10 (very 
excellent). I'll convert 'quality' as a factor variable to help me in 
exploration. Before that, I'll remove the the 'X' variable because it isn't 
needed in my analysis. 

```{r echo=FALSE}
# To delete column "x"
white_wine <- white_wine[,-1]

# To convert quality column from integer to factor
white_wine[,'quality'] <- factor(white_wine[,'quality'])

# print the data frame
str(white_wine)
```

```{r echo=FALSE, data_summary}
summary(white_wine)
```

After deleting 'X' variable (which is just an index) our dataset consists of 12 
variables, with 4898 observations.

**Hint 1:** The variables which I think they are the most variables affecting 
the 'quality' will have the blue histograms in this section.

**Hint 2:** In this analysis, I calculated outliers which are bigger than 
"Q3 + (1.5(IQR))" and smaller than "Q1 - (1.5 (IQR))" and I will use this 
calculations next in all variables. Now, let's start.

```{r echo=FALSE, quality}
ggplot(aes(x = quality), data = white_wine) +
    geom_bar(color = 'blue', fill = '#000080') 
```

```{r echo=FALSE, quality_summary}
table(white_wine$quality)
```

**The main variable 'quality':** I used bar plot because it is suitable for 
visualizing 'quality' as Factor/Categorical variable. Quality score is between 
(0 and 10) and the values range between 3 as minimum and 9 as maximum score. 
There are no observations in scores 1, 2 and 10. Most wine is in 6 score and 
the is least in 9 score. The wine evaluations (quality scores) are concentrated 
in three columns (5, 6 and 7). 

```{r echo=FALSE, fixed.acidity}
ggplot(aes(x = fixed.acidity), data = white_wine) +
    geom_histogram(binwidth = 0.1, color = 'black', fill = '#C0C0C0') 
```

```{r echo=FALSE, fixed.acidity_summary}
summary(white_wine$fixed.acidity)
```

**Fixed acidity (Tartaric acid - g/dm^3):** Most acids involved in wine are 
fixed or nonvolatile (do not evaporate readily). The fixed acidity follows 
a normal distribution with mean 6.855 (the median 6.800 g/dm^3) and some 
outliers found over 8.8 and under 4.8. 75% of wine is below 7.3 g/dm^3.

```{r echo=FALSE, customized_fixed.acidity}
ggplot(aes(x = fixed.acidity), data = white_wine) +
    geom_histogram(binwidth = 0.1, color = 'black', fill = '#C0C0C0') +
    coord_cartesian(xlim = c(5, 9)) +
    geom_vline(xintercept=mean(white_wine$fixed.acidity), color="red") +
    geom_vline(xintercept=median(white_wine$fixed.acidity), color="blue")
```

The plot got more readable by removing some outliers in the right side of the 
plot. Now, The histogram looks like approximately normal distribution.

```{r echo=FALSE, volatile.acidity}
ggplot(aes(x = volatile.acidity), data = white_wine) +
    geom_histogram(binwidth = 0.01, color = 'black', fill = '#C0C0C0')
```

```{r echo=FALSE, volatile.acidity_summary}
summary(white_wine$volatile.acidity)
```

**Volatile acidity (Acetic acid - g/dm^3):** The amount of acetic acid in wine, 
which at too high levels can **lead to an unpleasant, vinegar taste.**
The volatile acidity has some outliers over 0.49 and under 0.05. The 75% of the 
data is below 0.32 g/dm^3.

```{r echo=FALSE, customized_volatile.acidity}
ggplot(aes(x = volatile.acidity), data = white_wine) +
    geom_histogram(binwidth = 0.01, color = 'black', fill = '#000080') +
    coord_cartesian(xlim = c(0.05, 0.49)) +
    geom_vline(xintercept=mean(white_wine$volatile.acidity), color="red") +
    geom_vline(xintercept=median(white_wine$volatile.acidity), color="blue")
```

The trimmed volatile acidity distribution appears right skewed with mean in 
0.28 g/dm^3.

```{r echo=FALSE, citric.acid}
ggplot(aes(x = citric.acid), data = white_wine) +
    geom_histogram(binwidth = .01, color = 'BLACK', fill = '#C0C0C0')
```

```{r echo=FALSE, citric.acid_summary}
summary(white_wine$citric.acid)
```

**Citric acid** found in small quantities (0, 1 g/dm^3), citric acid can **add                  freshness and flavor to wines.** Outliers are found over 0.57 g/dm^3 and under
0.09 g/dm^3 and unusual spikes in 0.5 and 0.75 g/dm^3. I'll zoom in.

```{r echo=FALSE, customized_citric.acid}
ggplot(aes(x = citric.acid), data = white_wine) +
    geom_histogram(binwidth = .01, color = 'white', fill = '#000080') +
    coord_cartesian(xlim = c(0.09, 0.5)) +
    geom_vline(xintercept=mean(white_wine$citric.acid), color="red") +
    geom_vline(xintercept=median(white_wine$citric.acid), color="blue")
```

The trimmed Citric acid looks approximately normal distribution with unusual 
peak in 0.49 g/dm^3, we can consider it as anomaly.

```{r echo=FALSE, pH}
ggplot(aes(x = pH), data = white_wine) +
    geom_histogram(binwidth = 0.01, color = 'black', fill = '#C0C0C0')
```

```{r echo=FALSE, pH_summary}
summary(white_wine$pH)
```

**pH:** describes how **acidic or basic a wine** is on a scale from 0 (very 
acidic) to 14 (very basic); most wines are between 3 - 4 on the pH scale. It 
means moderately acidic. Outliers are found over 3.565 and under 2.805 and 
maximum acidic level is 3.82. 

```{r echo=FALSE, customized_pH}
ggplot(aes(x = pH), data = white_wine) +
    geom_histogram(binwidth = 0.01, color = 'black', fill = '#000080') +
    coord_cartesian(xlim = c(2.8, 3.57)) +
    geom_vline(xintercept=mean(white_wine$pH), color="red") +
    geom_vline(xintercept=median(white_wine$pH), color="blue")
```

The histogram is skewed a little bit to right but it is nearly normal with mean 
3.188 and median 3.180.

```{r echo=FALSE, residual.sugar}
ggplot(aes(x = residual.sugar), data = white_wine) +
    geom_histogram(binwidth = 1, color = 'black', fill = '#C0C0C0') 
```

**Residual sugar:** The amount of **sugar remaining** after fermentation stops, 
it's rare to find wines with less than 1 g/liter (g/dm^3) and wines with greater 
than 45 g/L are considered sweet. Most wine lie in 1 and 20 g/dm^3. Outliers 
are found in 23, 26, 32 and 66. let's remove outliers ..

```{r echo=FALSE, residual.sugar_summary}
summary(white_wine$residual.sugar)
```

```{r echo=FALSE, customized_residual.sugar}
ggplot(aes(x = residual.sugar), data = white_wine) +
    geom_histogram(binwidth = 1, color = 'black', fill = '#000080') +
    coord_cartesian(xlim = c(0, 22.2)) +
    geom_vline(xintercept=mean(white_wine$residual.sugar), color="red") +
    geom_vline(xintercept=median(white_wine$residual.sugar), color="blue")
```

Residual sugar is skewed to the right, with mean 6.39 g/dm^3 and the median 5.2 
g/dm^3 which resisted outliers to represent true medium value. Most wine are in
1 and 2 g/dm^3.
Here I used 'xlim' to remove outliers but it doesn't show clearly the difference 
between the peaks so I'll transform the x axis of to log10 to explore the data 
by zooming it and trim the long tail.

```{r echo=FALSE, log_transformed_residual.sugar}
ggplot(aes(x = residual.sugar), data = white_wine) +
    geom_histogram(binwidth = 0.05, color = 'black', fill = '#000080') +
    scale_x_log10() +
    geom_vline(xintercept=mean(white_wine$residual.sugar), color="red") +
    geom_vline(xintercept=median(white_wine$residual.sugar), color="blue")
```

The transformed distribution appears bi-modal with the peaks at (around 2 and 
again around 8 g/dm^3). 

Here I used the log10 of residual sugar rather than using limiting to focus on 
the distribution of the small values of data (from 0 to about 20).

```{r echo=FALSE, alcohol}
ggplot(aes(x = alcohol), data = white_wine) +
    geom_histogram(binwidth = 0.1, color = 'black', fill = '#C0C0C0') 
```

```{r echo=FALSE, alcohol_summary}
summary(white_wine$alcohol)
```

**Alcohol:** the percent alcohol content of the wine. The range is between 8% as 
minimum and 14.2% as maximum.

```{r echo=FALSE, customized_alcohol}
ggplot(aes(x = alcohol), data = white_wine) +
    geom_histogram(binwidth = 0.03, color = 'black', fill = '#000080') +
    scale_x_sqrt() +
    geom_vline(xintercept=mean(white_wine$alcohol), color="red") +
    geom_vline(xintercept=median(white_wine$alcohol), color="blue")
```

The histogram is clearly skewed to right with mean 10.51% and median 10.4 %.

```{r echo=FALSE, density}
ggplot(aes(x = density), data = white_wine) +
    geom_histogram(binwidth = 0.001, color = 'black', fill = '#C0C0C0')
```

```{r echo=FALSE, density_summary}
summary(white_wine$density)
```

**The density of water** is close to that of water depending on the percent 
alcohol and sugar content. Some outliers in 1.01 and 1.039 g/cm^3 
(g / cm^3 = 1000 g/dm^3) using Google. Let's trim it.

```{r echo=FALSE, customized_density}
ggplot(aes(x = density), data = white_wine) +
    geom_histogram(binwidth = 0.0003, color = 'black', fill = '#000080') +
    coord_cartesian(xlim = c(0.986, 1.0027)) +
    geom_vline(xintercept=mean(white_wine$density), color="red") +
    geom_vline(xintercept=median(white_wine$density), color="blue")
```

The shape appears as slightly right skewed distribution with mean equal 0.994 
g/cm^3 and median 0.9937 g/cm^3.  

```{r echo=FALSE, free.sulfur.dioxide}
ggplot(aes(x = free.sulfur.dioxide), data = white_wine) +
    geom_histogram(binwidth = 5, color = 'black', fill = '#C0C0C0') 
```

```{r echo=FALSE, free.sulfur.dioxide_summary}
summary(white_wine$free.sulfur.dioxide)
```

**Free sulfur dioxide:** The free form of SO2 exists in equilibrium between 
molecular SO2 (as a dissolved gas) and bisulfite ion; **it advances microbial                     growth and the oxidation of wine.** It has many outliers from 85 to 289 mg/dm^3 
as a maximum value. Let's trim it.

```{r echo=FALSE, customized_free.sulfur.dioxide}
ggplot(aes(x = free.sulfur.dioxide), data = white_wine) +
    geom_histogram(binwidth = 1, color = 'black', fill = '#C0C0C0') +
    coord_cartesian(xlim = c(0, 80)) +
    geom_vline(xintercept=mean(white_wine$free.sulfur.dioxide), color="red") +
    geom_vline(xintercept=median(white_wine$free.sulfur.dioxide), color="blue")
```

It is slightly right skewed but approximately normal and one peak at 29 mg/dm^3 
(0.029 g/dm^3).

```{r echo=FALSE, total.sulfur.dioxide}
ggplot(aes(x = total.sulfur.dioxide), data = white_wine) +
    geom_histogram(binwidth = 5, color = 'black', fill = '#C0C0C0') 
```

```{r echo=FALSE, total.sulfur.dioxide_summary}
summary(white_wine$total.sulfur.dioxide)
```

```{r echo=FALSE, total.sulfur.dioxide_summary<50}
summary(white_wine$total.sulfur.dioxide < 50)
```

**Total sulfur dioxide** is amount of free and bound forms of S02; in low 
concentrations, SO2 is mostly undetectable in wine, but at free SO2
concentrations over 50 ppm (1 ppm = 1 mg/dm^3) using Google, SO2 becomes evident 
in the nose and taste of wine. 
I found 46 of wine less than 50 ppm, So I asked, if SO2 becomes clear in smell 
and taste, is this good quality or bad quality of wine?
It has many outliers greater than 255 and below 19 mg/dm^3, with maximum 440 and 
minimum 9 mg/dm^3.

**Hint:** (According to Wikipedia article) In recent studies white wine (which 
contains high levels of Sulfur dioxide) was indicated to have some cardiovascular
benefits and this makes me excited to investigate the effect of Sulfur dioxide 
levels (which is considered **a health criterion**) on 'quality' of wine. 
We'll see.

```{r echo=FALSE, customized_total.sulfur.dioxide}
ggplot(aes(x = total.sulfur.dioxide), data = white_wine) +
    geom_histogram(binwidth = 5, color = 'black', fill = '#000080') +
    coord_cartesian(xlim = c(19, 260)) +
    geom_vline(xintercept=mean(white_wine$total.sulfur.dioxide), color="red") +
    geom_vline(xintercept=median(white_wine$total.sulfur.dioxide), color="blue")
```

After trimming the plot of total sulfur dioxide, it appears slightly right 
skewed with the mean at 138.4 and median at 134 mg/dm^3.

```{r echo=FALSE, chlorides}
ggplot(aes(x = chlorides), data = white_wine) +
    geom_histogram(binwidth = 0.001, color = 'black', fill = '#C0C0C0') 
```

```{r echo=FALSE, chlorides_summary}
summary(white_wine$chlorides)
```

**Chlorides (Sodium Chloride)** is the amount of **salt** in the wine (g/dm^3). 
It has many outliers greater than 0.071 and less than 0.015, with maximum 0.35 
and minimum 0.009 g/dm^3. Let's trim it.

```{r echo=FALSE, customized_chlorides}
ggplot(aes(x = chlorides), data = white_wine) +
    geom_histogram(binwidth = 0.001, color = 'black', fill = '#C0C0C0') +
    coord_cartesian(xlim = c(0.01, 0.071)) +
    geom_vline(xintercept=mean(white_wine$chlorides), color="red") +
    geom_vline(xintercept=median(white_wine$chlorides), color="blue")
```

After adjusting x axis from 0.01 to 0.071, Chlorides follows approximately 
normal but a little right skewed distribution, with max two values at 0.036 and 
0.044 g/dm^3 and the median is 0.043 g/dm^3.

```{r echo=FALSE, sulphates}
ggplot(aes(x = sulphates), data = white_wine) +
    geom_histogram(binwidth = 0.01, color = 'black', fill = '#C0C0C0') 
```

```{r echo=FALSE, sulphates_summary}
summary(white_wine$sulphates)
```

**Sulphates (potassium sulphate):** A wine additive which can contribute to 
sulfur dioxide gas (S02) levels, which acts as an antimicrobial and antioxidant. 
The distribution range between 0.22 and 1.08 g/dm^3, and it has outliers upper 
0.76 and below 0.2. I'll trim it.

```{r echo=FALSE, customized_sulphates}
ggplot(aes(x = sulphates), data = white_wine) +
    geom_histogram(binwidth = 0.01, color = 'black', fill = '#C0C0C0') +
    coord_cartesian(xlim = c(0.2, 0.7)) +
    geom_vline(xintercept=mean(white_wine$sulphates), color="red") +
    geom_vline(xintercept=median(white_wine$sulphates), color="blue")
```

The histogram is slightly right skewed with mean 0.48 and median 0.47 g/dm^3.

# Univariate Analysis

### What is the structure of your dataset?

This dataset contains 4,898 records of white wine data with 12 variables, 
11 input variables based on physicochemical properties which affect wine quality
and 1 output variable which is the quality of the wine. At least 3 wine experts 
rated the quality of each wine, providing a rating between 0 (very bad) and 10 
(very excellent).

### What is/are the main feature(s) of interest in your dataset?

I think 'quality' is the main feature in this dataset because it is a criterion 
of wine goodness. I want to check all the other variables combinations, which of 
them are affecting the 'quality' the most. Quality scores (between 0 and 10) 
values range (in the dataset) between 3 as minimum and 9 as maximum score. There 
are no observations in scores 1, 2 and 10. Most frequent 'quality' score is in 6 
score and the least frequent is in 9 score. 
Since the most frequent 'quality' scores are in the three columns (5, 6 and 7)
in the middle, does this indicate that the 'quality' evaluation is based on the 
experts opinions of taste and smell? Or on a specific technical basis depending 
on the physical or chemical properties of the wine?. So, this is what I will 
investigate in the next section.

### What other features in the dataset do you think will help support your                  investigation into your feature(s) of interest?

I want to investigate the two variables (residual sugar and alcohol) because I 
read about them that they are affecting the 'quality' of wine and I observed 
that they have clearly different distributions than the other variables.
Also I want to do more investigations on some variables (Total Sulfur dioxide, 
Density, Volatile acidity, Citric acid and pH) which I think they are important 
for their influence (smell and taste) on wine quality.

### Did you create any new variables from existing variables in the dataset?

I didn't need to create any new variables.

### Of the features you investigated, were there any unusual distributions? Did                      you perform any operations on the data to tidy, adjust, or change the form of                        the data? If so, why did you do this?

I converted 'quality' column as a factor variable to help me in exploration.
When plotting histograms of the variables, I applied the 'xlim' to limit the 
lower and upper x-axis value to remove outliers and to focus on the main values 
for better visualizations and to see clearer distributions of these variables. 
Also, I log-transformed the right skewed 'Residual Sugar' distribution to 
zoom in the small values of data.

# Bivariate Plots Section

```{r echo=FALSE, correlation_matrix}
M <- cor(white_wine[1:11])
# lower triangular matrix
corrplot(M, type = "lower")
```

The correlation matrix shows interesting relations between (residual.sugar vs 
density) and (alcohol vs density), also between (residual sugar and alcohol). 
This gives me a hint that I should try (residual.sugar, alcohol & density) as 
a combination in Multi-variate analysis. I will also check acidity related 
variables (pH, fixed.acidity and volatile.acidity) vs the other variables to 
determine the over all effect of acidity on the other variables.

```{r echo=FALSE, density_vs_residual.sugar}
ggplot(aes(x = density, y = residual.sugar), data = white_wine) +
    geom_point(alpha = .5, position = position_jitter(), size = 1) +
    coord_cartesian(xlim = c(0.985, 1.01), ylim = c(0, 20)) + 
    geom_smooth(method = 'lm', color = 'yellow') 
```

```{r echo=FALSE, corr_density_vs_residual.sugar}
cor.test(white_wine$density, white_wine$residual.sugar)
```

I found strong positive correlation (+0.84) between density and residual.sugar 

```{r echo=FALSE, density_vs_alcohol}
ggplot(aes(x = density, y = alcohol), data = white_wine) +
    geom_point(alpha = .5, position = position_jitter(), size = 1) +
    coord_cartesian(xlim = c(0.985, 1.02), ylim = c(7.5, 15)) + 
    geom_smooth(method = 'lm', color = 'blue') 
```

```{r echo=FALSE, corr_density_vs_alcohol}
cor.test(white_wine$density, white_wine$alcohol)
```

Strong positive correlation (about 0.8) is between (density and residual.sugar)
but Strong negative correlation about 0.8 is between density and alcohol, It 
makes sense, considering the fermentation process of wine, the sugar is consumed
to produce alcohol. And since the residual sugar is more denser than alcohol 
(shown above). The inverse relation presented and negative moderate correlation
(-0.45), as shown next.

```{r echo=FALSE, residual.sugar_vs_alcohol}
ggplot(aes(x = residual.sugar, y = alcohol), data = white_wine) +
    geom_point(alpha = .5, position = position_jitter(), size = 1) +
    coord_cartesian(xlim = c(0, 35)) +
    geom_smooth(method = 'lm', color = 'blue', se = FALSE) 
```

```{r echo=FALSE, corr_residual.sugar_vs_alcohol}
cor.test(white_wine$residual.sugar, white_wine$alcohol)
```

Now, what is the relation between each of these three variables (residual.sugar, 
alcohol and density) with 'quality'

```{r echo=FALSE, quality_vs_residual.sugar}
ggplot(aes(x = quality, y = residual.sugar), data = white_wine) +
    geom_jitter(alpha = .5, color = 'orange')  +
    geom_boxplot(color = 'dark blue', fill = 'dark cyan') +
    coord_cartesian(ylim = c(0, 15)) +
    stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8, 
                 size = 4)
```

```{r echo=FALSE, quality_summary2}
table(white_wine$quality)
```

Residual sugar is the variable with the negative correlation with 'quality' 
except grades 3 and 4 may be because as shown in the last cell the number of 
data samples in (3, 4, 8 & 9) are too small compared with (5, 6 & 7) **so** 
**(5, 6 & 7) may be good representatives of the main trend of these scores.** 

```{r echo=FALSE, quality_vs_alcohol}
ggplot(aes(x = quality, y = alcohol), data = white_wine) +
    geom_jitter(alpha = .5, color = 'orange')  +
    geom_boxplot(color = 'dark blue', fill = 'dark cyan') +
    coord_cartesian(ylim = c(7.5, 14))  +
    stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8, 
                 size = 4)
```

```{r echo=FALSE, alcohol_vs_quality_summary}
by(white_wine$alcohol, white_wine$quality, summary)
```

Alcohol is the variable with positive correlation with 'quality'. Wines with 
scores 3 and 4 are going against the trend may be because there are too small 
number of data examples in scores (3 and 4) so it is hard to capture a trend 
from them.

```{r echo=FALSE, quality_vs_density}
ggplot(aes(x = quality, y = density), data = white_wine) +
    geom_jitter(alpha = 0.5, color = 'orange')  +
    geom_boxplot(color = 'dark blue', fill = 'dark cyan') +
    coord_cartesian(ylim = c(0.989, 1.0)) +
    stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8, 
                 size = 4)
```

Here clear negative correlation between 'quality' and 'density' is shown. This 
correlation is expected as (density & residual.sugar) are strongly positively 
correlated and both of them is negatively correlated with 'quality'.

```{r echo=FALSE, alcohol_vs_total.sulfur.dioxide}
ggplot(aes(x = alcohol, y = total.sulfur.dioxide), data = white_wine) +
    geom_point(position = position_jitter(), size = 1.5) +
    geom_smooth(method = 'lm', color ='blue') 
```

```{r echo=FALSE, corr_alcohol_vs_total.sulfur.dioxide}
cor.test(white_wine$total.sulfur.dioxide, white_wine$alcohol)
```

Alcohol has bad effects on Liver. However, sulfur dioxide has cardiovascular 
benefits. They have negative correlation, **what a strange coincidence!**

```{r echo=FALSE, quality_vs_total.sulfur.dioxide}
ggplot(aes(x = quality, y = total.sulfur.dioxide), data = white_wine) +
    geom_jitter(alpha = .5, color = 'orange')  +
    geom_boxplot(color = 'dark blue', fill = 'dark cyan') +
    coord_cartesian(ylim = c(50, 250)) +
    stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8, 
                 size = 4)
```

```{r echo=FALSE, total.sulfur.dioxide_vs_quality_summary}
by(white_wine$total.sulfur.dioxide, white_wine$quality, summary)
```

Here a slight correlation can be seen, Overall quality increases as total sulfur 
dioxide decreases, except scores 3 and 4 with least wines. It is strange that as 
the sulfur dioxide decreases the quality increases which **makes me sure that** 
**quality does not depend on scientific measurements** because if it depends on 
scientific measurements it may make the quality increase with sulfur dioxide 
(or at least like the rate of increase in alcohol with quality) because sulfur 
dioxide has some health benefits.

```{r echo=FALSE, fixed.acidity_vs_pH}
ggplot(aes(x = fixed.acidity, y = pH), data = white_wine) +
    geom_point(alpha = 0.5, position = position_jitter()) +
    coord_cartesian(ylim = c(2.5, 4)) +
    geom_smooth(method = 'lm', color = 'blue', se = FALSE) 
```

```{r echo=FALSE, corr_fixed.acidity_vs_pH}
cor.test(white_wine$fixed.acidity, white_wine$pH)
```

The fixed acidity concentration shows negative moderate correlation with pH.

```{r echo=FALSE, quality_vs_volatile.acidity}

ggplot(aes(x = quality, y = volatile.acidity), data = white_wine) +
    geom_jitter(alpha = .5, color = 'orange')  +
    geom_boxplot(color = 'dark blue', fill = 'dark cyan') +
    coord_cartesian(ylim = c(0.1, 0.5)) +
    stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8, 
                 size = 4)
```

The wine quality increases when volatile acidity decreases, this is what I 
expected because when (volatile acidity) increases it can lead to an unpleasant,
vinegar taste.

```{r echo=FALSE, quality_vs_fixed.acidity}

ggplot(aes(x = quality, y = fixed.acidity), data = white_wine) +
    geom_jitter(alpha = .5, color = 'orange')  +
    geom_boxplot(color = 'dark blue', fill = 'dark cyan') +
    coord_cartesian(ylim = c(6, 9)) +
    stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8, 
                 size = 4)
```

Also, better quality wines seem to have smaller fixed acidifies on average. This 
is what I expected too because if pH decreases, fixed acidity decreases. 

```{r echo=FALSE, quality_vs_pH}
ggplot(aes(x = quality, y = pH), data = white_wine) +
    geom_jitter(alpha = .5, color = 'orange')  +
    geom_boxplot(color = 'dark blue', fill = 'dark cyan') +
    coord_cartesian(ylim = c(3, 3.5)) +
    stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8, 
                 size = 4)
```

Except for wines with quality score 3, the median pH increases (Acidity 
decreases) along with quality score. So, from the last three visualizations I 
conclude that as acidity decreases 'quality' increases.

```{r echo=FALSE, quality_chlorides}
ggplot(aes(x = quality, y = chlorides), data = white_wine) +
    geom_jitter(alpha = .5, color = 'orange')  +
    geom_boxplot(color = 'dark blue', fill = 'dark cyan') +
    coord_cartesian(ylim = c(0.015, 0.075)) +
    stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8, 
                 size = 4) 
```

The wine quality increases when chlorides decrease.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the                 investigation. How did the feature(s) of interest vary with other features in                        the dataset?

I  found 'quality' has positive moderate correlation with 'alcohol' and I found 
also 'quality' has negative correlation with 'residual.sugar', **I found these** 
**results make sense because fermentation process is the conversion of sugar to** 
**alcohol so sugar and alcohol have negative correlation.**
'quality' has negative correlation with density.
I found also that 'quality' has negative correlation with total sulfur dioxide
which **makes me observe that 'quality' rating is not based on the healthy** 
**components (like sulfur dioxide).**
There are 3 variables to represent acidity quantity (pH, fixed acidity and 
volatile acidity) both fixed acidity and volatile acidity have negative 
correlation with 'quality' and pH has positive correlation (acidity decreases) 
with 'quality', so the 3 variables say the quality has negative correlation with 
acidity.
Finally I found also 'quality' has negative correlation with chlorides.

### Did you observe any interesting relationships between the other features                        (not the main feature(s) of interest)?

One of the exciting relationships that I found is alcohol has negative 
correlation with total sulfur dioxide which **may mean as the healthy components** 
**(sulfur dioxide) increases the harmful components (alcohol) decreases.** 
Residual sugar has negative correlation with alcohol due to fermentation process 
as explained before.
pH has negative correlation with fixed acidity and volatile acidity and this is 
expected because **pH scale decreases as acidity increases.**
Also I found density has negative correlation with alcohol and positive 
correlation with sugar and this is expected as sugar is always decreases as 
alcohol increases (Fermentation).

```{r echo=FALSE, quality_table}
table(white_wine$quality)
```

**Hint:** The 'quality' (3, 4, 8 and 9) scores (as shown in the last cell) has 
so small number of data points compared to the other scores which gives 
inaccurate trending results in almost all the boxplots of the last section and 
also they became too much sensitive to outliers so not showing the right trend 
of data.

### What was the strongest relationship you found?

- density has strong negative correlation with alcohol (-0.7801376)
- density has strong positive correlation with sugar ( 0.8389665)

# Multivariate Plots Section

```{r echo=FALSE, residual.sugar_vs_density_by_alcohol}
ggplot(aes(x = residual.sugar, y = density, color = alcohol), 
       data = white_wine) +
    coord_cartesian(xlim = c(0, 23), ylim = c(0.985, 1.01)) +
    geom_point() +
    stat_smooth(method = 'lm', color = 'yellow', se = FALSE) +
    ggtitle('Residual Sugar vs Density By Alcohol')
```

I started with this plot because it shows the relation between the main 3 
features affecting the wine quality (having the biggest correlation coefficients
in the Bi-variate section) so we can conclude that these 3 features are the main
focus of the experts to rate the wine quality. The visualization shows two 
observations: the first is **sweeter wine has more density**, the second is 
**wine with the same sweetness has larger percentage of alcohol when density**
**increases.**And the reverse is correct.

```{r echo=FALSE, fixed.acidity_vs_pH_by_citric.acid}
ggplot(aes(x = fixed.acidity, y = pH, color = citric.acid), 
       data = white_wine) +
    geom_point(position = position_jitter(), size = 2) +
    stat_smooth(method = 'lm', color = 'black', se = FALSE) +
    ggtitle('Fixed Acidity vs pH By Citric Acid')
```

The plot shows a moderate negative correlation between (pH and fixed.acidity) 
and this is expected because pH is a measure for acidity (pH decreases as 
acidity increases). I can justify the gathering of the data points in this 
specific area and not spread because **most wines are between 3-4 on the pH**
**scale.** Finally I observed that citric acid has no clear trend which means 
that it has no significant effect on pH of wine.

```{r echo=FALSE, Alcohol_vs_Total.Sulfur.Dioxide_by_Quality}
ggplot(aes(x = alcohol, y = total.sulfur.dioxide, color = quality), 
       data = white_wine) +
    geom_point(position = position_jitter(), size = 2) +
    stat_smooth(method = 'lm', color = 'black', se = FALSE) +
    ggtitle('Alcohol vs Total Sulfur Dioxide By quality')
```

```{r echo=FALSE, alcohol_vs_Total.Sulfur.Dioxide_by_quality_2}
ggplot(aes(x = alcohol, y = total.sulfur.dioxide), data = white_wine) +
    geom_point(color = 'blue', position = position_jitter(), size = 1) +
    stat_smooth(method = 'lm', color = 'black', se = FALSE) +
    facet_wrap(~ quality)
```

The 2 plots show a moderate negative correlation (-0.45) between (Total sulfur 
dioxide and alcohol). Generally, I noticed the spread of the green and blue 
colors among the data points which represent quality scores (5,6 and 7) and this
observation makes sense as in the quality variable distribution big number of 
data points are in scores (5,6 and 7) and if the other scores (3, 4, 8 and 9) 
have also a big number of data points we could have captured a more clear trend 
of quality scores than the shown plot. Finally, the captured trend of quality 
score is as the alcohol increases the quality score increases but total sulfur 
dioxide decreases.  

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the                   investigation. Were there features that strengthened each other in terms of                       looking at your feature(s) of interest?

From (Residual Sugar vs density by alcohol) plot I observed strong positive 
correlation between (sugar and density) while alcohol decreases as density 
increases. Also in (Acidity variables) I observed a moderate negative 
correlation between (pH and fixed.acidity). Finally, a moderate negative 
correlation between (Total sulfur dioxide and alcohol) and positive correlation 
between (alcohol and quality).

### Were there any interesting or surprising interactions between features?

The negative correlation between (Total sulfur dioxide and alcohol) which shows 
the conflict and the trade-off between the healthy components (Total sulfur 
dioxide) and the harmful components (alcohol).

------

# Final Plots and Summary

### Plot One

```{r echo=FALSE, Plot_One_quality}
ggplot(aes(x = quality), data = white_wine) +
    geom_bar(color = 'black', fill = 'blue') +
    ggtitle('Quality of White Wine') +
    xlab('Quality Scores (ranges from 0 to 10)') +
    ylab('Number of White Wine')
```

### Description One

The quality variable is the main feature of concern in my analysis. The 
distribution of quality scores is normal. Not distributing the data uniformly 
makes some difficulties in data analysis and not capturing clear trends. Also 
there are only about 5000 data points representing the white wine data which are
not large enough data points to get a strong analysis. 

### Plot Two
```{r echo=FALSE, Plot_Two_quality_vs_alcohol}
p1 <- ggplot(aes(x = quality, y = alcohol), data = white_wine) +
    geom_jitter(alpha = .1, color = 'Darkorange')  +
    geom_boxplot(color = 'black', fill = 'dark cyan')+
    coord_cartesian(ylim = c(8, 14)) +
    stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8, 
               size = 4) +
    ggtitle('Quality Scores vs Alcohol') +
    xlab('Quality Scores (ranges from 0 to 10)') +
    ylab('Alcohol (% of Wine Volume)')

p2 <- ggplot(aes(x = quality, y = total.sulfur.dioxide), data = white_wine) +
    geom_jitter(alpha = .1, color ='DarkGreen')  +
    geom_boxplot(color = 'black', fill = 'dark cyan')+
    coord_cartesian(ylim = c(50, 250)) +
    stat_summary(fun.y = "mean", geom = "point", color = "red", shape = 8, 
               size = 4) +
    ggtitle('Quality Scores vs Total Sulfur Dioxide') +
    xlab('Quality Scores (ranges from 0 to 10)') +
    ylab('Total Sulfur Dioxide (mg/dm^3)')

grid.arrange(p1, p2, ncol = 2)
```

### Description Two

Here, this is the most exciting result to me which is the trade-off between 
Alcohol (as a harmful component) and Total Sulfur Dioxide (as a healthy 
component) which gives me impression that there must be a balanced quantities 
for both of them to reach the maximum possible benefit and the minimum possible 
harmful effects. Therefore, according to my readings the weekly healthy dose of 
drinking wine is 6 cups per week which can minimize the harmful effects as much 
as possible and to try to benefit from it. 

### Plot Three
```{r echo=FALSE, Plot_Three_residual.sugar_vs_density_by_alcohol}
ggplot(aes(x = residual.sugar, y = density, color = alcohol), 
       data = white_wine) +
    coord_cartesian(xlim = c(0, 22), ylim = c(0.985, 1.01)) +
    geom_point(position = position_jitter(), size = 1.5) +
    stat_smooth(method = 'lm', color = 'yellow') +
    ggtitle('Residual Sugar vs Density By Alcohol') +
    xlab('Residual Sugar (g/dm^3)') +
    ylab('Density (g/cm^3)')
```

### Description Three

From this plot I can conclude that as sugar increases alcohol decreases which 
shows that some people prefer high levels of sugar and the other people prefer 
low levels of sugar and according to my readings the manufacturers take this 
point in consideration during fermentation process so it needs accurate tools. 

------

# Reflection

Beside data analysis I made a lot of exploration and reading about the wine 
components and types to be able to describe the relations between variables and 
to determine the main features of concern during analysis that was because in my
culture drinking wine is not popular among people and also it is known that wine
has harmful effects more than its benefits, so I was excited to check this by 
myself through data analysis. 
The surprising result I concluded was that as the harmful components (alcohol) 
increases the healthy components (total sulfur dioxide) decreases.
The struggles that I faced is the relatively low number of data points, so 
capturing trends among the variables was not an easy task. In the future I am so
interested to repeat this analysis if there are dataset available with large 
number of data points.

# References:

Readings:
White wine: https://en.wikipedia.org/wiki/White_wine
https://www.spiritswinecellar.com/components-of-wine/

Coding:
https://stackoverflow.com/questions/34617883/how-to-remove-multiple-columns-in-
r-dataframe
https://stackoverflow.com/questions/31165015/how-to-convert-integer-to-
factor-in-r
https://stackoverflow.com/questions/6557977/how-do-i-add-the-mean-value-to- a-histogram-in-r
https://rstudio-pubs-static.s3.amazonaws.com/240657_5157ff98e8204c358b2118fa69162e18.html
