
White wine quality by exploring physicochemical properties data
by Raafat Saleh
===============

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Loading all packages that I'll use in my analysis in this code chunk.
library(ggplot2)
library(ggthemes)
library(dplyr)

```

```{r echo=FALSE, Load_the_Data}
# Load the Data
white_wine <- read.csv("wineQualityWhites.csv", sep = ",")
```

# Introduction

   The dataset is related to **"White Wine"** of the Portuguese "Vinho Verde" 
wine. For more details, consult: http://www.vinhoverde.pt/en/ or the reference [Cortez et al., 2009]. Due to privacy and logistic issues, only physicochemical (inputs) and sensory (the output) variables are available (e.g. there is no data about grape types, wine brand, wine selling price, etc.). Also, we are not sure
if all input variables are relevant. So it could be interesting to test feature selection methods. 

- Number of Instances (obs.): 4898. 
- Number of Attributes (variables): 11 + output attribute + serial No (x var.)

# Univariate Data Exploring and Plots Section

> **Tip**: In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

```{r echo=FALSE, dimensions}
dim(white_wine)
```

Our dataset consists of 12 variables, with 4898 observations.

```{r echo=FALSE, string}
str(white_wine)
```

I notice 11 variables are numerics, on quantifying the chemical properties of 
the wine, except quality variable is integer. At least 3 wine experts rated the quality of each wine, providing a rating between 0 (very bad) and 10 (very excellent). # I'll convert quality as a factor variable to help me for 
exploration. Befor that, I'll remove the the x variable because it isn't needed 
in my analysis.

```{r echo=True, delete x column, convert quality from integer to factor}
# To delete column "x"
white_wine <- white_wine[,-1]

# To convert quality column from integer to factor
#white_wine[,'quality'] <- factor(white_wine[,'quality'])

str(white_wine)
```

Ref: 
https://stackoverflow.com/questions/34617883/how-to-remove-multiple-columns-in-
r-dataframe
https://stackoverflow.com/questions/31165015/how-to-convert-integer-to-
factor-in-r

```{r echo=FALSE, data_summary}
summary(white_wine)
```

> **Tip**: Make sure that you leave a blank line between the start / end of
each code block and the end / start of your Markdown text so that it is
formatted nicely in the knitted text. Note as well that text on consecutive
lines is treated as a single space. Make sure you have a blank line between
your paragraphs so that they too are formatted for easy readability.

```{r echo=FALSE, quality}
ggplot(aes(x = quality), data = white_wine) +
    geom_histogram() 
```

```{r echo=FALSE, quality_summary}
summary(white_wine$quality)
```

Quality score (between 0 and 10) values range between 3 as minimum and 9 as maximum level, with 5.88 is mean and 6 is median (approximatly appears unimodal normal distribution with center in 6). 75% of quality score below 6. 

```{r echo=FALSE, fixed.acidity}
ggplot(aes(x = fixed.acidity), data = white_wine) +
    geom_histogram(binwidth = 0.1) +
    geom_vline(xintercept=mean(white_wine$fixed.acidity), color="red") +
    geom_vline(xintercept=median(white_wine$fixed.acidity), color="blue")
```

Ref: https://stackoverflow.com/questions/6557977/how-do-i-add-the-mean-value-to-
a-histogram-in-r

```{r echo=FALSE, fixed.acidity_summary}
summary(white_wine$fixed.acidity)
```

Fixed acidity (tartaric acid - g/dm^3) most acids involved with wine or fixed or nonvolatile (do not evaporate readily). The fixed acidity follows a normal distribution with mean 6.855 (the median 6.800 g/dm^3) and some outliers over 14 and under 4 (3.800). The 75% of the data is below 7.3 g/dm^3.

To make the plot more readable and remove some outliers in the right side of the plot, 99 percentile is used to limit the upper bound of the xlimit. The histogram now look more a normal distribution.

```{r echo=FALSE, customized fixed.acidity}
ggplot(aes(x = fixed.acidity), data = white_wine) +
    geom_histogram(binwidth = 0.1) +
    scale_x_continuous(limits = c(4, quantile(white_wine$fixed.acidity, 0.99))) +
    geom_vline(xintercept=mean(white_wine$fixed.acidity), color="red") +
    geom_vline(xintercept=median(white_wine$fixed.acidity), color="blue")
```

```{r echo=FALSE, volatile.acidity}
ggplot(aes(x = volatile.acidity), data = white_wine) +
    geom_histogram() +
    geom_vline(xintercept=mean(white_wine$volatile.acidity), color="red") +
    geom_vline(xintercept=median(white_wine$volatile.acidity), color="blue")
```

```{r echo=FALSE, volatile.acidity_summary}
summary(white_wine$volatile.acidity)
```

```{r echo=FALSE, volatile.acidity}
ggplot(aes(x = volatile.acidity), data = white_wine) +
    geom_histogram(binwidth = 0.01) +
    scale_x_continuous(limits = c(0.08, quantile(
                       white_wine$volatile.acidity, 0.95))) +
    geom_vline(xintercept=mean(white_wine$volatile.acidity), color="red") +
    geom_vline(xintercept=median(white_wine$volatile.acidity), color="blue")
```
Volatile acidity (acetic acid - g/dm^3) the amount of acetic acid in wine, which at too high of levels can lead to an unpleasant, vinegar taste. The customized volatile acidity distribution appears normal with the peaking around 0.28 g/dm^3

```{r echo=FALSE, citric.acid}
ggplot(aes(x = citric.acid), data = white_wine) +
    geom_histogram(binwidth = .02) +
    geom_vline(xintercept=mean(white_wine$citric.acid), color="red") +
    geom_vline(xintercept=median(white_wine$citric.acid), color="blue")
```

```{r echo=FALSE, citric.acid_summary}
summary(white_wine$citric.acid)
```

```{r echo=FALSE, citric.acid}
ggplot(aes(x = citric.acid), data = white_wine) +
    geom_histogram(binwidth = .02) +
    scale_x_continuous(limits = c(0, quantile(white_wine$volatile.acidity, 
                                              0.99))) +
    geom_vline(xintercept=mean(white_wine$citric.acid), color="red") +
    geom_vline(xintercept=median(white_wine$citric.acid), color="blue")
```

```{r echo=FALSE, residual.sugar}
ggplot(aes(x = residual.sugar), data = white_wine) +
    geom_histogram() +
    geom_vline(xintercept=mean(white_wine$residual.sugar), color="red") +
    geom_vline(xintercept=median(white_wine$residual.sugar), color="blue")
```



```{r echo=FALSE, residual.sugar_summary}
summary(white_wine$residual.sugar)
```

```{r echo=FALSE, residual.sugar2}
ggplot(aes(x = residual.sugar), data = white_wine) +
    geom_histogram(binwidth = 0.5) +
    geom_vline(xintercept=mean(white_wine$residual.sugar), color="red") +
    geom_vline(xintercept=median(white_wine$residual.sugar), color="blue")
```

```{r echo=FALSE, chlorides}
ggplot(aes(x = chlorides), data = white_wine) +
    geom_histogram() +
    geom_vline(xintercept=mean(white_wine$chlorides), color="red") +
    geom_vline(xintercept=median(white_wine$chlorides), color="blue")
```

```{r echo=FALSE, chlorides_summary}
summary(white_wine$chlorides)
```

```{r echo=FALSE, free.sulfur.dioxide}
ggplot(aes(x = free.sulfur.dioxide), data = white_wine) +
    geom_histogram() +
    geom_vline(xintercept=mean(white_wine$free.sulfur.dioxide), color="red") +
    geom_vline(xintercept=median(white_wine$free.sulfur.dioxide), color="blue")
```

```{r echo=FALSE, free.sulfur.dioxide_summary}
summary(white_wine$free.sulfur.dioxide)
```

```{r echo=FALSE, total.sulfur.dioxide}
ggplot(aes(x = total.sulfur.dioxide), data = white_wine) +
    geom_histogram() +
    geom_vline(xintercept=mean(white_wine$total.sulfur.dioxide), color="red") +
    geom_vline(xintercept=median(white_wine$total.sulfur.dioxide), color="blue")
```

```{r echo=FALSE, total.sulfur.dioxide_summary}
summary(white_wine$total.sulfur.dioxide)
```

```{r echo=FALSE, density}
ggplot(aes(x = density), data = white_wine) +
    geom_histogram() +
    geom_vline(xintercept=mean(white_wine$density), color="red") +
    geom_vline(xintercept=median(white_wine$density), color="blue")
```

```{r echo=FALSE, density_summary}
summary(white_wine$density)
```

```{r echo=FALSE, pH}
ggplot(aes(x = pH), data = white_wine) +
    geom_histogram() +
    geom_vline(xintercept=mean(white_wine$pH), color="red") +
    geom_vline(xintercept=median(white_wine$pH), color="blue")
```

```{r echo=FALSE, pH_summary}
summary(white_wine$pH)
```

```{r echo=FALSE, sulphates}
ggplot(aes(x = sulphates), data = white_wine) +
    geom_histogram() +
    geom_vline(xintercept=mean(white_wine$sulphates), color="red") +
    geom_vline(xintercept=median(white_wine$sulphates), color="blue")
```

```{r echo=FALSE, sulphates_summary}
summary(white_wine$sulphates)
```

```{r echo=FALSE, alcohol}
ggplot(aes(x = alcohol), data = white_wine) +
    geom_histogram() +
    geom_vline(xintercept=mean(white_wine$alcohol), color="red") +
    geom_vline(xintercept=median(white_wine$alcohol), color="blue")
```

```{r echo=FALSE, alcohol_summary}
summary(white_wine$alcohol)
```

# Univariate Analysis

> **Tip**: Now that you've completed your univariate explorations, it's time to
reflect on and summarize what you've found. Use the questions below to help you
gather your observations and add your own if you have other thoughts!

### What is the structure of your dataset?
4898 observations,  13 variables
### What is/are the main feature(s) of interest in your dataset?

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \ 
of the data? If so, why did you do this?


# Bivariate Plots Section

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

pairs(white_wine[1:11])

```

```{r echo=FALSE, alcohol_residual.sugar}
ggplot(aes(x = alcohol, y = residual.sugar), data = white_wine) +
    geom_point()
```

```{r echo=FALSE, corr_alcohol_residual.sugar}
cor.test(white_wine$alcohol, white_wine$residual.sugar)
```

```{r echo=FALSE, alcohole_sulphates}
ggplot(aes(x = alcohol, y = sulphates), data = white_wine) +
    geom_point()
```

```{r echo=FALSE, corr_alcohole_sulphates}
cor.test(white_wine$alcohol, white_wine$sulphates)
```

```{r echo=FALSE, alcohole_pH}
ggplot(aes(x = alcohol, y = pH), data = white_wine) +
    geom_point()
```

```{r echo=FALSE, corr_alcohole_pH}
cor.test(white_wine$alcohol, white_wine$pH)
```

```{r echo=FALSE, alcohole_density}
ggplot(aes(x = alcohol, y = density), data = white_wine) +
    geom_point()
```

```{r echo=FALSE, corr_alcohole_density}
cor.test(white_wine$alcohol, white_wine$density)
```

```{r echo=FALSE, alcohole_volatile.acidity}
ggplot(aes(x = alcohol, y = volatile.acidity), data = white_wine) +
    geom_point()
```

```{r echo=FALSE, corr_alcohole_volatile.acidity}
cor.test(white_wine$alcohol, white_wine$volatile.acidity)
```

```{r echo=FALSE, alcohole_fixed.acidity}
ggplot(aes(x = alcohol, y = fixed.acidity), data = white_wine) +
    geom_point()
```

```{r echo=FALSE, corr_alcohole_fixed.acidity}
cor.test(white_wine$alcohol, white_wine$fixed.acidity)
```

```{r echo=FALSE, alcohole_citric.acid}
ggplot(aes(x = alcohol, y = citric.acid), data = white_wine) +
    geom_point()
```

```{r echo=FALSE, corr_alcohole_citric.acid}
cor.test(white_wine$alcohol, white_wine$citric.acid)
```

```{r echo=FALSE, alcohole_total.sulfur.dioxide}
ggplot(aes(x = alcohol, y = total.sulfur.dioxide), data = white_wine) +
    geom_point()
```

```{r echo=FALSE, corr_alcohole_total.sulfur.dioxide}
cor.test(white_wine$alcohol, white_wine$total.sulfur.dioxide)
```

```{r echo=FALSE, alcohole_chlorides}
ggplot(aes(x = alcohol, y = quality), data = white_wine) +
    geom_point()
```

```{r echo=FALSE, corr_alcohole_chlorides}
cor.test(white_wine$alcohol, white_wine$chlorides)
```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!

